# Data aquisition and understanding

## Description

```{r libraries and scripts, message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse)
library(corrplot)
library(car)
library(tidymodels)
library(mice)
library(forecast)

# script with visual functions
source("scripts/eda_functions.r")

# import data
unbalanced_data <- read_csv("./data/train_full.csv") %>% select(-...1)
balanced_data <- read_csv("./data/train_balanced.csv") %>% select(-...1)

# convert categorical and dummy to factor variable
unbalanced_data <- unbalanced_data %>% 
  mutate_at(vars(
    !contains(c("duration", "view_qty", "unique_product_qty","unique_browse_designer_qty"))
    ), ~ as.factor(.))

balanced_data <- balanced_data %>% 
  mutate_at(vars(
    !contains(c("duration", "view_qty", "unique_product_qty","unique_browse_designer_qty"))
    ), ~ as.factor(.))

```

Two datasets were made available for analysis. One with a balanced class for the target variable and the other raw. This analysis focus on the full dataset. 

From the available description the available columns follow into one of the following categories:

1. Irrelevant for the project;

2. Target or dependent Variable: what will be studied;

3. Customer Behavior related Features: potential features which reflect how users behaved on each session and can explain the target variable;

4. Customer Journey related Features: potential features contain information on how uses interacted with the on-line shop;

5. Client Segment related Features: potential features which provide information on how a user might be classified from a marketing perspective;

6. Device and Geography related Features: provide information regarding location and access device used by a user.

```{r data structure, message=FALSE, warning=FALSE}

str(unbalanced_data)

```

## Irrelevant or technical information

`Session_id` is a unique key which sole purpose is to uniquely identify each sessions and contains no exploratory power, therefore it will be removed from future analysis.

```{r}

# test if a single session can have more than one row
length( unique(unbalanced_data$session_id) ) == nrow(unbalanced_data)

```
## Target Variable: To buy or not to buy ...

**Boolean variable** which assumes the values of 1 when a transaction happened on the session and zero other wise. Its the objective variable of this project.

The available dataset with full data presents a severe class imbalance towards `not buy` which can and will Bias model output (a model assuming that nobody buys on each session is already "92% of the times right"). This bias is corrected on the balanced dataset at the expanse of some changes on 

```{r target variable, echo=FALSE, include=FALSE}

# unbalanced eda
unbalanced_bar <- ggplot(unbalanced_data, aes(x = as.numeric(bought))) +
    geom_bar(fill="darkblue", color="black", aes(y = ..prop..), stat = "count") +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Frequency"
    ) +
  scale_y_continuous(labels=scales::percent) +
  geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.5)

unbalanced_summary <- unbalanced_data %>% select(bought) %>% summary()

unbalanced_grid <- grid.arrange(unbalanced_bar, tableGrob(unbalanced_summary), 
                                nrow = 1, top=textGrob("Unbalanced dataset"))

# balanced eda
balanced_bar <- ggplot(balanced_data, aes(x = as.numeric(bought) )) +
    geom_bar(fill="darkblue", color="black", aes(y = ..prop..), stat = "count") +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Frequency"
    ) +
  scale_y_continuous(labels=scales::percent) +
  geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.5)

balanced_summary <- balanced_data %>% select(bought) %>% summary()

balanced_grid <- grid.arrange(balanced_bar, tableGrob(balanced_summary), nrow = 1, top=textGrob("Unbalanced dataset"))

```

```{r target unbalanced grid, echo=FALSE}

grid.draw(unbalanced_grid) 

```

```{r target balanced grid}

grid.draw(balanced_grid)

```


