## Customer Behavior Features

In this section we will explore the variables available using both unbalanced and balanced data. In special we will focus on the following questions:

- What type of variation occurs within my variables?
- What type of covariation occurs between my variables?

```{r, message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse)
library(corrplot)
library(car)
library(tidymodels)
library(mice)
library(forecast)

# script with visual functions
source("scripts/eda_functions.r")

# import data
unbalanced_data <- read_csv("./data/train_full.csv") %>% select(-...1)
balanced_data <- read_csv("./data/train_balanced.csv") %>% select(-...1)

# convert categorical and dummy to factor variable
unbalanced_data <- unbalanced_data %>% 
  mutate_at(vars(
    !contains(c("duration", "view_qty", "unique_product_qty",
                "unique_browse_designer_qty", "unique_browse_category_qty"))
    ), ~ as.factor(.))

balanced_data <- balanced_data %>% 
  mutate_at(vars(
    !contains(c("duration", "view_qty", "unique_product_qty",
                "unique_browse_designer_qty", "unique_browse_category_qty"))
    ), ~ as.factor(.))

```

```{r unbalanced data eda plots, include=FALSE, include=FALSE}

continuous <- unbalanced_data %>% 
  select(duration) %>% 
  names() %>% 
  set_names()
 
discrete <- unbalanced_data %>% 
  select(view_qty, unique_product_qty, unique_browse_designer_qty) %>% 
  names() %>% 
  set_names()

categorical <- unbalanced_data %>% 
  select(-duration, -view_qty, -unique_product_qty, 
         -unique_browse_designer_qty, -session_id, -bought) %>% 
  names() %>% 
  set_names()

eda_cont <- map(continuous, ~eda_continuous(unbalanced_data, .x))

eda_dis <- map(discrete, ~eda_discrete(unbalanced_data, .x))

eda_cat <- map(categorical, ~eda_categorical(unbalanced_data, .x))

eda_unbalanced <- c(eda_cont, eda_dis, eda_cat)
rm(eda_cont,eda_dis, eda_cat)

```

```{r balanced data eda plots, balanced data, include=FALSE, include=FALSE}

continuous <- balanced_data %>% 
  select(duration) %>% 
  names() %>% 
  set_names()

discrete <- balanced_data %>% 
  select(view_qty, unique_product_qty, unique_browse_designer_qty) %>% 
  names() %>% 
  set_names()

categorical <- balanced_data %>% 
  select(-duration, -view_qty, -unique_product_qty, 
         -unique_browse_designer_qty, -session_id, -bought) %>% 
  names() %>% 
  set_names()

eda_cont <- map(continuous, ~eda_continuous(balanced_data, .x))

eda_dis <- map(discrete, ~eda_discrete(balanced_data, .x))

eda_cat <- map(categorical, ~eda_categorical(balanced_data, .x))

eda_balanced <- c(eda_cont, eda_dis, eda_cat)
rm(eda_cont,eda_dis, eda_cat)

```

```{r correlation plot, correlation plot, include=FALSE}

M <- cor(unbalanced_data %>% 
           select_if(is.numeric))
corrplot(M, type = "upper", method = "number", diag = FALSE, 
         order = 'hclust', addrect = 2)

```

```{r Chi-squared hipothesis test, message=FALSE, warning=FALSE, include=FALSE}

unbalanced_chi <- unbalanced_data %>% 
  select_if(is.factor) %>% 
  map(function(x) chisq.test(x, unbalanced_data$bought))

balanced_chi <- balanced_data %>% 
  select_if(is.factor) %>% 
  map(function(x) chisq.test(x, balanced_data$bought))

unbalanced_chi_numeric <- unbalanced_data %>% 
  select_if(is.numeric) %>% 
  map(function(x) chisq.test(x, as.factor(unbalanced_data$bought)))

balanced_chi_numeric <- balanced_data %>% 
  select_if(is.numeric) %>% 
  map(function(x) chisq.test(x, as.factor(balanced_data$bought)))

```

### Duration

Continuous variable representing the time a user was on a particular sessions. Values measured in seconds. No specifics are given regarding the concept of session used. The market standard is to limit session duration to an engagement window between 25 to 30 minutes. Above that time cap, a new session would start (even if the user is still navigating the site).
It will assumed that no such cap exist and therefore extreme values are **not due to technical issues on web metrics**.

```{r plot duration unbalanced, echo=FALSE}

grid.draw(eda_unbalanced$duration) 

```

```{r plot duration balanced, echo=FALSE}

grid.draw(eda_balanced$duration) 

```

Distribution information shows a severely left skewed distribution which is confirmed by the behavior of the qq_plot. This is present on both raw and balanced data. The max duration of `r max(unbalanced_data$duration)` means that at least one session was `{r} max(unbalanced_data$duration)/(60 * 60)` hours long, which, given what was said regarding session cap won't be considered as an outlier due to error.

Data shows that visit duration is a long tail with a great majority of users staying just a fraction but a long line of users (power users?) which stay for longer. This distribution does not allow for a good analysis of outliers of the distribution given its nature and might prove to be tricky during modeling due to its non Gaussian nature. 

```{r log transform duration, echo=TRUE}

  density <- ggplot(unbalanced_data, aes(x = duration)) +
    geom_density(color="darkblue") +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Density"
    )
  
  boxplot <- ggplot(unbalanced_data, aes(x = 1, y = duration)) + 
    geom_boxplot() + 
    theme_masterDS()+
    labs(
      x = "",
      y="",
      title = "Boxplot"
    )   

  density_log <- ggplot(unbalanced_data, aes(x = log(duration))) +
    geom_density(color="darkblue") +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Density (log transform)"
    )
  
  boxplot_log <- ggplot(unbalanced_data, aes(x = 1, y = log(duration))) + 
    geom_boxplot() + 
    theme_masterDS()+
    labs(
      x = "",
      y="",
      title = "Boxplot (log transform)"
    ) 

grid.arrange(density, boxplot, density_log, boxplot_log,  layout_matrix = rbind(c(1,2),c(3,4)))

```

**As a preprocess step this feature will be log transformed.** No more complex Box-Cox transformation will be used given that good results can already be achieved with the current transformation. No missing values were found. 

```{r duration vs bought, echo=FALSE}

ggplot(unbalanced_data, aes(x = bought, y = log(duration))) + 
    geom_boxplot() + 
    theme_masterDS()+
    labs(
      x = "",
      y="",
      title = "Boxplot by Conversion",
      caption = "Data from unbalanced dataset.Log transformed"
    ) 

```

```{r duration vs bought balanced, echo=FALSE}

ggplot(balanced_data, aes(x = bought, y = log(duration))) + 
    geom_boxplot() + 
    theme_masterDS()+
    labs(
      x = "",
      y="",
      title = "Boxplot by Conversion",
      caption = "Data from balanced dataset. Log transformed"
    ) 

```

Ploted against the target variable it seems to imply a positive relationship between duration and conversion. From the sample sessions that converted had on median a higher duration. 


### Page View Quantity

Discrete variable measuring the number of views during a session. 

```{r page view unbalanced, echo=FALSE}

grid.draw(eda_unbalanced$view_qty)

```


```{r page view balanced, echo=FALSE}

grid.draw(eda_balanced$view_qty)

```

Distribution information shows a severely left skewed distribution which is confirmed by the behavior of the qq_plot. This is present on both raw and balanced data. The max duration is of `r max(unbalanced_data$view_qty)`. 

Despite is distribution the data shows a median of 10 page views which suggest an interesting engagement for a e-commerce provided other indicators regarding navigation are solid (if a user is jumping between pages because they don't load properly, an increase on page views might no be beneficial).

Given the available information we cannot conclude that outliers due to errors exist. In an attempt to normalize the distribution a Box-Cox transformation will be used with a Optimized Lambda. 

```{r view_qty box cox unbalanced, echo=FALSE, message=TRUE, warning=FALSE}
lambda <- BoxCox.lambda(unbalanced_data$view_qty, method = "guerrero") 

histogram_log <- ggplot(unbalanced_data, aes(x = bcPower(view_qty, lambda = lambda))) +
    geom_histogram(fill="darkblue", color="black") +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Histogram"
    )
  
boxplot_log <- ggplot(unbalanced_data, aes(x = 1, y = bcPower(view_qty, lambda = lambda))) +
    geom_boxplot() +
    theme_masterDS()+
    labs(
      x = "",
      y="",
      title = "Boxplot",
      caption = "unbalanced data"
    )

grid.arrange(histogram_log, boxplot_log, nrow = 1)

```


```{r view_qty box cox balanced, echo=FALSE, message=TRUE, warning=FALSE}

lambda <- BoxCox.lambda(balanced_data$view_qty, method = "guerrero") 

histogram_log <- ggplot(balanced_data, aes(x = bcPower(view_qty, lambda = lambda))) +
    geom_histogram(fill="darkblue", color="black") +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Histogram"
    )
  
boxplot_log <- ggplot(balanced_data, aes(x = 1, y = bcPower(view_qty, lambda = lambda))) +
    geom_boxplot() +
    theme_masterDS()+
    labs(
      x = "",
      y="",
      title = "Boxplot",
      caption = "unbalanced data"
    )

grid.arrange(histogram_log, boxplot_log, nrow = 1)

```
With this transformation it is now visible 2 segments of sessions which deserve attention, sessions with zero views or just 1 view. 

The information available provides no information whatsoever regardingt definition of view. The standard (defined by google) is the first hit at log to be a pageview. Therefore, technically all sessions would have a view. Using the same approach, a 1 view session is technically called a **bounce**. Once again assuming Googles standard (it owns 97% of the analytics market ...) a bounce is a session with just a starting hit but not information regarding log out (duration is a difference based calculation). 

On a standard scenario the approach would be to remove all observations with a zero pageviews (technical errors) and session with 1 page view (if not other information regarding duration is available). In the scenario in analysis, given the available information we can't assume zero page views to be a technical error and therefore will filter the **dataset for observations with zero or one page views and a session duration smaller than 2 seconds.**


```{r unbalanced no outlier, echo=FALSE, message=TRUE, warning=FALSE}

lambda <- BoxCox.lambda(balanced_data$view_qty, method = "guerrero") 

data <- unbalanced_data %>% 
  filter( !((view_qty <= 1) & (duration <= 2)))

histogram_log <- ggplot(data, aes(x = bcPower(view_qty, lambda = lambda))) +
    geom_histogram(fill="darkblue", color="black", bins = 15) +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Histogram"
    )
  
boxplot_log <- ggplot(data, aes(x = 1, y = bcPower(view_qty, lambda = lambda))) +
    geom_boxplot() +
    theme_masterDS()+
    labs(
      x = "",
      y="",
      title = "Boxplot",
      caption = "balanced data"
    )

grid.arrange(histogram_log, boxplot_log, nrow = 1)

```


```{r balanced no outlier, echo=FALSE, message=TRUE, warning=FALSE}

lambda <- BoxCox.lambda(balanced_data$view_qty, method = "guerrero") 

data <- balanced_data %>% 
  filter( !((view_qty <= 1) & (duration <= 2)))

histogram_log <- ggplot(data, aes(x = bcPower(view_qty, lambda = lambda))) +
    geom_histogram(fill="darkblue", color="black", bins = 15) +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Histogram"
    )
  
boxplot_log <- ggplot(data, aes(x = 1, y = bcPower(view_qty, lambda = lambda))) +
    geom_boxplot() +
    theme_masterDS()+
    labs(
      x = "",
      y="",
      title = "Boxplot",
      caption = "balanced data"
    )

grid.arrange(histogram_log, boxplot_log, nrow = 1)

```

The balanced data to be used with modeling is already close to normal distribution.


```{r pageview vs bought balanced, echo=FALSE}

ggplot(data, aes(x = bought, y = log(view_qty))) + 
    geom_boxplot() + 
    theme_masterDS()+
    labs(
      x = "",
      y="",
      title = "Boxplot by Conversion",
      caption = "Data from balanced dataset. Log transformed"
    ) 

```
Compared by conversion the data seems to suggest that sessions with conversion had a higher number of page views.

### Product Pages unique page views

Discrete variable measuring the visits product related pages had. 

```{r eda_unique_product unbalanced}

grid.draw(eda_unbalanced$unique_product_qty)

```


```{r eda_unique_product balanced}

grid.draw(eda_balanced$unique_product_qty)

```

The initial graphic analysis shows a right skewed distribution similar to was seen on before on Pageviews. Nonetheless, in this scenario zero page views is a acceptable outcome. When the target class is taken into consideration it seems to sugest a teneous impact albeit its true effects might be hidden by the distribution skew. 

```{r pageviews vs product, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = unbalanced_data, 
       aes(x = unique_product_qty, y = view_qty, 
           shape = bought, color = bought)) +
  geom_point(alpha = 0.5, position = position_jitter()) +
  stat_smooth(method = lm, level = 0.95, se = TRUE, size = 0.5) +
  scale_shape_manual(values = c(1,2)) +
  scale_colour_brewer(palette = "Set1") +
  theme_masterDS() +
    labs(
      x = "Pageviews product pages",
      y = "Pageviews" ,
      title = "Pageviews and product pageviews",
      caption = "Data from unbalanced dataset"
    ) 

```

Ploted together target, pageviews and product pageviews it becomes more evident that the relationship between target and product pageviews is not very visable but the Chi test lets us renouce the null hypothesis. 

```{r echo=FALSE}

unbalanced_chi_numeric$unique_product_qty

```

On the other hand, there seems to have a strong relation positive relation between pageviews and pagevies product. This may suggest an interaction between both features. Given the nature of this feature we will explore the impact of converting into a ratio using page views as denominator since product page view is in fact a subgroup of pageviews.



```{r pageviews vs product ratio, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = unbalanced_data, 
       aes(x = unique_product_qty / view_qty, y = view_qty, 
           shape = bought, color = bought)) +
  geom_point(alpha = 0.5, position = position_jitter()) +
  scale_shape_manual(values = c(1,2)) +
  scale_colour_brewer(palette = "Set1") +
  theme_masterDS() +
    labs(
      x = "Pageviews product pages",
      y = "Pageviews" ,
      title = "Pageviews and product/pageview ",
      caption = "Data from unbalanced dataset"
    ) 

```

```{r unbalanced with ratio product, echo=FALSE, message=TRUE, warning=FALSE}

data <- unbalanced_data %>% 
  filter( !((view_qty <= 1) & (duration <= 2)))

histogram_log <- ggplot(data, aes(x = log (unique_product_qty / view_qty ))) +
    geom_histogram(fill="darkblue", color="black", bins = 15) +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Histogram"
    )
  
boxplot_log <- ggplot(data, aes(x = 1, y = log(unique_product_qty / view_qty ))) +
    geom_boxplot() +
    theme_masterDS() +
    labs(
      x = "",
      y="",
      title = "Boxplot",
      caption = "unbalanced data"
    )

grid.arrange(histogram_log, boxplot_log, nrow = 1)

```
```{r balanced with ratio product, echo=FALSE, message=TRUE, warning=FALSE}

data <- balanced_data %>% 
  filter( !((view_qty <= 1) & (duration <= 2)))

histogram_log <- ggplot(data, aes(x = log(unique_product_qty / view_qty ))) +
    geom_histogram(fill="darkblue", color="black", bins = 15) +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Histogram"
    )
  
boxplot_log <- ggplot(data, aes(x = 1, y = log( unique_product_qty / view_qty ))) +
    geom_boxplot() +
    theme_masterDS() +
    labs(
      x = "",
      y="",
      title = "Boxplot",
      caption = "balanced data"
    )

grid.arrange(histogram_log, boxplot_log, nrow = 1)

```
This transformation provides more insights into the behaviour of the average user indicating that on average 25% of page navigation is done on product related pages (from and into). Simultaneously it smooths the skewed effect reported earlier

```{r ratio product view vs bought balanced, echo=FALSE}

ggplot(data, aes(x = bought, y = log(unique_product_qty / view_qty ))) + 
    geom_boxplot() + 
    theme_masterDS()+
    labs(
      x = "",
      y="",
      title = "Boxplot by Conversion",
      caption = "Data from balanced dataset. Log transformed"
    ) 

```

If the e-commecer has a more transactional focus it might not come as surprice the above plot and it seems to imply that user might already know what they want to buy prior to visit.

Similar transformation will be done to the subsquent variables-

### Designer Browsing unique page views

Discrete variable measuring the number of Designers researched.


```{r}

grid.draw(eda_unbalanced$unique_browse_designer_qty)

```


```{r}

grid.draw(eda_balanced$unique_browse_designer_qty)

```

Similar approach done before


```{r pageviews vs designer, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = unbalanced_data, 
       aes(x = unique_browse_designer_qty, y = view_qty, 
           shape = bought, color = bought)) +
  geom_point(alpha = 0.5, position = position_jitter()) +
  stat_smooth(method = lm, level = 0.95, se = TRUE, size = 0.5) +
  scale_shape_manual(values = c(1,2)) +
  scale_colour_brewer(palette = "Set1") +
  theme_masterDS() +
    labs(
      x = "Pageviews product pages",
      y = "Pageviews" ,
      title = "Pageviews and designer pageviews",
      caption = "Data from unbalanced dataset"
    ) 

```

```{r unbalanced with ratio designer, echo=FALSE, message=TRUE, warning=FALSE}

data <- unbalanced_data %>% 
  filter( !((view_qty <= 1) & (duration <= 2)))

histogram_log <- ggplot(data, aes(x = log(unique_browse_designer_qty / view_qty ))) +
    geom_histogram(fill="darkblue", color="black", bins = 15) +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Histogram"
    )
  
boxplot_log <- ggplot(data, aes(x = 1, y = log(unique_browse_designer_qty / view_qty ))) +
    geom_boxplot() +
    theme_masterDS() +
    labs(
      x = "",
      y="",
      title = "Boxplot",
      caption = "unbalanced data"
    )

grid.arrange(histogram_log, boxplot_log, nrow = 1)

```

```{r balanced with ratio designer, echo=FALSE, message=TRUE, warning=FALSE}

data <- balanced_data %>% 
  filter( !((view_qty <= 1) & (duration <= 2)))

histogram_log <- ggplot(data, aes(x = log(unique_browse_designer_qty / view_qty ))) +
    geom_histogram(fill="darkblue", color="black", bins = 15) +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Histogram"
    )
  
boxplot_log <- ggplot(data, aes(x = 1, y = log(unique_browse_designer_qty / view_qty ))) +
    geom_boxplot() +
    theme_masterDS() +
    labs(
      x = "",
      y="",
      title = "Boxplot",
      caption = "balanced data"
    )

grid.arrange(histogram_log, boxplot_log, nrow = 1)

```

This transformation provides more insights into the behaviour of the average user indicating that on average 25% of page navigation is done on product related pages (from and into). Simultaneously it smooths the skewed effect reported earlier

```{r ratio designer view vs bought balanced, echo=FALSE}

ggplot(data, aes(x = bought, y = log(unique_browse_designer_qty / view_qty ))) + 
    geom_boxplot() + 
    theme_masterDS() +
    labs(
      x = "",
      y="",
      title = "Boxplot by Conversion",
      caption = "Data from balanced dataset. Log transformed"
    ) 

```

### Products categories unique id

Discrete variable measuring the visits Designer related pages had.


```{r}

grid.draw(eda_unbalanced$unique_browse_category_qty)

```


```{r}

grid.draw(eda_balanced$unique_browse_category_qty)

```

Similar approach done before


```{r pageviews vs category, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = unbalanced_data, 
       aes(x = unique_browse_category_qty, y =  view_qty , 
           shape = bought, color = bought)) +
  geom_point(alpha = 0.5, position = position_jitter()) +
  stat_smooth(method = lm, level = 0.95, se = TRUE, size = 0.5) +
  scale_shape_manual(values = c(1,2)) +
  scale_colour_brewer(palette = "Set1") +
  theme_masterDS() +
    labs(
      x = "Pageviews product pages",
      y = "Pageviews" ,
      title = "Pageviews and designer pageviews",
      caption = "Data from unbalanced dataset"
    ) 

```

```{r echo=FALSE}

unbalanced_chi$unique_browse_category_qty


```



```{r unbalanced with ratio category, echo=FALSE, message=TRUE, warning=FALSE}

data <- unbalanced_data %>% 
  filter( !((view_qty <= 1) & (duration <= 2)))

histogram_log <- ggplot(data, aes(x = log(unique_browse_category_qty / view_qty ))) +
    geom_histogram(fill="darkblue", color="black", bins = 15) +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Histogram"
    )
  
boxplot_log <- ggplot(data, aes(x = 1, y = log(unique_browse_category_qty  / view_qty ))) +
    geom_boxplot() +
    theme_masterDS() +
    labs(
      x = "",
      y="",
      title = "Boxplot",
      caption = "unbalanced data"
    )

grid.arrange(histogram_log, boxplot_log, nrow = 1)

```

```{r balanced with ratio category, echo=FALSE, message=TRUE, warning=FALSE}

data <- balanced_data %>% 
  filter( !((view_qty <= 1) & (duration <= 2)))

histogram_log <- ggplot(data, aes(x = log(unique_browse_category_qty / view_qty ))) +
    geom_histogram(fill="darkblue", color="black", bins = 15) +
    theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Histogram"
    )
  
boxplot_log <- ggplot(data, aes(x = 1, y = log(unique_browse_category_qty / view_qty ))) +
    geom_boxplot() +
    theme_masterDS() +
    labs(
      x = "",
      y="",
      title = "Boxplot",
      caption = "balanced data"
    )

grid.arrange(histogram_log, boxplot_log, nrow = 1)

```

This transformation provides more insights into the behaviour of the average user indicating that on average 25% of page navigation is done on product related pages (from and into). Simultaneously it smooths the skewed effect reported earlier

```{r ratio category view vs bought balanced, echo=FALSE}

ggplot(data, aes(x = bought, y = log(unique_browse_category_qty / view_qty ))) + 
    geom_boxplot() + 
    theme_masterDS()+
    labs(
      x = "",
      y="",
      title = "Boxplot by Conversion",
      caption = "Data from balanced dataset. Log transformed"
    ) 

```


### Potential ideas for feature engineering which were not implemented

- The available data could synthesized using a non-supervised cluster algorithm which would help identify subsets of behaviors to be used during modeling. This approach has the advantage of possible reduce 4 variables into one.

















