## Client Segment related Features

```{r, message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse)
library(corrplot)
library(car)
library(tidymodels)
library(mice)
library(forecast)

# script with visual functions
source("scripts/eda_functions.r")

# import data
unbalanced_data <- read_csv("./data/train_full.csv") %>% select(-...1)
balanced_data <- read_csv("./data/train_balanced.csv") %>% select(-...1)

# convert categorical and dummy to factor variable
unbalanced_data <- unbalanced_data %>% 
  mutate_at(vars(
    !contains(c("duration", "view_qty", "unique_product_qty",
                "unique_browse_designer_qty", "unique_browse_category_qty"))
    ), ~ as.factor(.))

balanced_data <- balanced_data %>% 
  mutate_at(vars(
    !contains(c("duration", "view_qty", "unique_product_qty",
                "unique_browse_designer_qty", "unique_browse_category_qty"))
    ), ~ as.factor(.))

```

```{r , include=FALSE, include=FALSE}

continuous <- unbalanced_data %>% 
  select(duration) %>% 
  names() %>% 
  set_names()
 
discrete <- unbalanced_data %>% 
  select(view_qty, unique_product_qty, unique_browse_designer_qty) %>% 
  names() %>% 
  set_names()

categorical <- unbalanced_data %>% 
  select(-duration, -view_qty, -unique_product_qty, 
         -unique_browse_designer_qty, -session_id, -bought) %>% 
  names() %>% 
  set_names()

eda_cont <- map(continuous, ~eda_continuous(unbalanced_data, .x))

eda_dis <- map(discrete, ~eda_discrete(unbalanced_data, .x))

eda_cat <- map(categorical, ~eda_categorical(unbalanced_data, .x))

eda_unbalanced <- c(eda_cont, eda_dis, eda_cat)
rm(eda_cont,eda_dis, eda_cat)

```

```{r , include=FALSE, include=FALSE}

continuous <- balanced_data %>% 
  select(duration) %>% 
  names() %>% 
  set_names()

discrete <- balanced_data %>% 
  select(view_qty, unique_product_qty, unique_browse_designer_qty) %>% 
  names() %>% 
  set_names()

categorical <- balanced_data %>% 
  select(-duration, -view_qty, -unique_product_qty, 
         -unique_browse_designer_qty, -session_id, -bought) %>% 
  names() %>% 
  set_names()

eda_cont <- map(continuous, ~eda_continuous(balanced_data, .x))

eda_dis <- map(discrete, ~eda_discrete(balanced_data, .x))

eda_cat <- map(categorical, ~eda_categorical(balanced_data, .x))

eda_balanced <- c(eda_cont, eda_dis, eda_cat)
rm(eda_cont,eda_dis, eda_cat)

```

```{r , include=FALSE}

M <- cor(unbalanced_data %>% 
           select_if(is.numeric))
corrplot(M, type = "upper", method = "number", diag = FALSE, 
         order = 'hclust', addrect = 2)

```

```{r , message=FALSE, warning=FALSE, include=FALSE}

unbalanced_chi <- unbalanced_data %>% 
  select_if(is.factor) %>% 
  map(function(x) chisq.test(x, unbalanced_data$bought))

balanced_chi <- balanced_data %>% 
  select_if(is.factor) %>% 
  map(function(x) chisq.test(x, balanced_data$bought))

unbalanced_chi_numeric <- unbalanced_data %>% 
  select_if(is.numeric) %>% 
  map(function(x) chisq.test(x, as.factor(unbalanced_data$bought)))

balanced_chi_numeric <- balanced_data %>% 
  select_if(is.numeric) %>% 
  map(function(x) chisq.test(x, as.factor(balanced_data$bought)))

```


### Customer type

Categorical variable with `r levels(unbalanced_data$customer_type)`.

```{r echo=FALSE}
grid.draw(eda_unbalanced$customer_type)
```

```{r echo=FALSE}
grid.draw(eda_balanced$customer_type)
```

The graphical analysis shows inbalance towards prospect clients. When plotted together with target class it suggests a potential relationship with customers have a higher propency to buy again (happy customers?) 

The results from the $\chi^2$ hypothesis test does not refuse the null hypothesis reinforcing the graphical analysis that a relationship might exist between this 2 variables that implies that recurrent customers buy more.

```{r}
unbalanced_chi["customer_type"]
```

```{r}
balanced_chi["customer_type"]
```

### Segment

Categorical variable with `r levels(unbalanced_data$segment)` levels. No information is provided regarding how this segment is generated and is highly unbalanced towards no segment.

```{r echo=FALSE}
grid.draw(eda_unbalanced$segment)
```

```{r}
unbalanced_chi["segment"]
```


```{r echo=FALSE}
grid.draw(eda_balanced$segment)
```

```{r}
balanced_chi["segment"]
```

Despite the low number of sessions made by users with a segment the ones which have seem to have a very high conversion rate despite their level. The number of levels on a categorical variable affects the model, and in this case the extra information doesn't seem to be bringing any new impute therefore, **it will be collapses into a new variable with just two levels, has segment or it has no segment.**


### Visitor type

Categorical variable with 2 levels representing if a given user is a new or recurring user. It differs from customer type because it focus on visits and not actual conversion, therefore a returning user can be a prospect.

```{r echo=FALSE}
grid.draw(eda_unbalanced$visitor_type)
```

```{r echo=FALSE}
unbalanced_chi["visitor_type"]
```

```{r echo=FALSE}
grid.draw(eda_balanced$visitor_type)
```

```{r echo=FALSE}
balanced_chi["visitor_type"]

```


This feature focus on the Visitors. The unbalanced data available shows almost a 50% split, situations that changes on the balanced dataset which has a inbalance towards returning visitors. The graphical analysis suggests a higher conversion rate for returning visitor that for new one, suggesting that continuous visits (engagment) plays a role in conversion.

The current dataset has information regarding customer type crossed with visitor type can provide us with interesting information


```{r}

ggplot(unbalanced_data, aes(x = visitor_type, fill = customer_type)) + 
  geom_bar(stat = "count") + 
  theme_masterDS() +
    labs(
      x = "",
      y = "",
      title = "Visitor Type x Custor Type"
    )

```

Has expected only a fraction of new visitors actually buy on the first session hinting that the conversion journey is longer than one session, meaning than several visits are needed before a first conversion. We don't have enough information to conclude about the number of sessions needed (journey lenght) and neither the session index given a time window (eg: the current session is the #3 in the last 28 days) which is know to have a impact on conversion.


### Is subscribed


```{r}

grid.draw(eda_unbalanced$is_subscribed)

```


```{r}

grid.draw(eda_balanced$is_subscribed)

```

For both datasets the class majority is NA. From the information given we cannot conclude that we can remove this feature from the model, and given the number of observations affected we will look for imputation alternatives.

If a particular variable is having more missing values that rest of the variables in the dataset, and, if by removing that one variable you can save many observations then is advisable to remove. But in this case, the fact a user is subscribed is a strong signal and one of the few (apart from segment) which provides insight about the companies marketing strategy. Removing this feature would mean removing signal regarding an ongoing engagement strategy. A concervative approach would be to consider all missing values as non subscribers. Given the context of this project an alternative root of statistical imputation will be used relying on a knn algorithm with a small number of neighbors (3 = 2 levels + 1)




