## Device and Geography related Features

```{r, message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse)
library(corrplot)
library(car)
library(tidymodels)
library(mice)
library(forecast)

# script with visual functions
source("scripts/eda_functions.r")

# import data
unbalanced_data <- read_csv("./data/train_full.csv") %>% select(-...1)
balanced_data <- read_csv("./data/train_balanced.csv") %>% select(-...1)

# convert categorical and dummy to factor variable
unbalanced_data <- unbalanced_data %>% 
  mutate_at(vars(
    !contains(c("duration", "view_qty", "unique_product_qty",
                "unique_browse_designer_qty", "unique_browse_category_qty"))
    ), ~ as.factor(.))

balanced_data <- balanced_data %>% 
  mutate_at(vars(
    !contains(c("duration", "view_qty", "unique_product_qty",
                "unique_browse_designer_qty", "unique_browse_category_qty"))
    ), ~ as.factor(.))

```

```{r unbalanced data eda plots journey, include=FALSE, include=FALSE}

continuous <- unbalanced_data %>% 
  select(duration) %>% 
  names() %>% 
  set_names()
 
discrete <- unbalanced_data %>% 
  select(view_qty, unique_product_qty, unique_browse_designer_qty) %>% 
  names() %>% 
  set_names()

categorical <- unbalanced_data %>% 
  select(-duration, -view_qty, -unique_product_qty, 
         -unique_browse_designer_qty, -session_id, -bought) %>% 
  names() %>% 
  set_names()

eda_cont <- map(continuous, ~eda_continuous(unbalanced_data, .x))

eda_dis <- map(discrete, ~eda_discrete(unbalanced_data, .x))

eda_cat <- map(categorical, ~eda_categorical(unbalanced_data, .x))

eda_unbalanced <- c(eda_cont, eda_dis, eda_cat)
rm(eda_cont,eda_dis, eda_cat)

```

```{r balanced data eda plots journey, balanced data, include=FALSE, include=FALSE}

continuous <- balanced_data %>% 
  select(duration) %>% 
  names() %>% 
  set_names()

discrete <- balanced_data %>% 
  select(view_qty, unique_product_qty, unique_browse_designer_qty) %>% 
  names() %>% 
  set_names()

categorical <- balanced_data %>% 
  select(-duration, -view_qty, -unique_product_qty, 
         -unique_browse_designer_qty, -session_id, -bought) %>% 
  names() %>% 
  set_names()

eda_cont <- map(continuous, ~eda_continuous(balanced_data, .x))

eda_dis <- map(discrete, ~eda_discrete(balanced_data, .x))

eda_cat <- map(categorical, ~eda_categorical(balanced_data, .x))

eda_balanced <- c(eda_cont, eda_dis, eda_cat)
rm(eda_cont,eda_dis, eda_cat)

```

```{r correlation plot journey, correlation plot, include=FALSE}

M <- cor(unbalanced_data %>% 
           select_if(is.numeric))
corrplot(M, type = "upper", method = "number", diag = FALSE, 
         order = 'hclust', addrect = 2)

```

```{r Chi-squared hipothesis test journey, message=FALSE, warning=FALSE, include=FALSE}

unbalanced_chi <- unbalanced_data %>% 
  select_if(is.factor) %>% 
  map(function(x) chisq.test(x, unbalanced_data$bought))

balanced_chi <- balanced_data %>% 
  select_if(is.factor) %>% 
  map(function(x) chisq.test(x, balanced_data$bought))

unbalanced_chi_numeric <- unbalanced_data %>% 
  select_if(is.numeric) %>% 
  map(function(x) chisq.test(x, as.factor(unbalanced_data$bought)))

balanced_chi_numeric <- balanced_data %>% 
  select_if(is.numeric) %>% 
  map(function(x) chisq.test(x, as.factor(balanced_data$bought)))

```


### Browser name

Categorical variable representing the browser used as source for each session. From the available information from the unbalanced data we have 52 different browsers present on this dataset. Most of this levels have only a small number of sessions and will be colapsed during the preprocess

```{r echo=FALSE}

unique(unbalanced_data$browser_name)

```


```{r echo=FALSE}

grid.draw(eda_unbalanced$browser_name)

```


```{r echo=FALSE}

grid.draw(eda_balanced$browser_name)

```

### Country

Categorical variable containing the country of origin for each session.Most of this levels have only a small number of sessions and will be colapsed during the preprocess

```{r echo=FALSE}

unique(unbalanced_data$country)

```


```{r echo=FALSE}

grid.draw(eda_unbalanced$country)

```


```{r echo=FALSE}

grid.draw(eda_balanced$country)

```

### Dealing with Missing Data

Somo missing data was identified and will be imputed using statistical imputation (refer to customer type analysis)

```{r echo=TRUE}
md.pattern(unbalanced_data, rotate.names = TRUE)
```

```{r echo=TRUE}
md.pattern(balanced_data,rotate.names = TRUE)
```



## Summary of findings

- Features *Device Group* and *Platform* showed a strong relationship between them. Given that *Platform* provides more information *Device Group will be removed*
- Features starting with *"has_"* refere to milestones on customer journey, therefore the need to include all of them should be taken in consideration. We will use PCA to merge this into one feature and experiment between the use of them individual or together,
- *Duration* feature is Right Skewed and contains observations which lack business sense. To avoid any impact on the model, specially on ones more sensible to variance, it will be log transformed and observations above 24hours will be removed
- Replace all *Missing values* on *country*, *browser_name* and *is_subscribed* using Nearest Neighbors imputation with a very a k of 3,
- Reduce the number of observations of class country and browser_name by compressing smaller ones into a generic class named others. This will be done because some models do not manage very well the fact that some dummy variables only have one level,
- Reduce the number of levels of Segment by re factoring into a class of has or not segment, 
- Specially for the use of Neural Networks and Supervisor Vector Machines, numeric variables will be normalized and scaled
- Converts *unique_browse_designer_qty, unique_product_qty, unique_browse_designer_qty, unique_browse_category_qty* to ratios over Total Pageviews (view_qty) which provides information on the amount of page views actually related to that category and subsquently log transform,
- Sessions with duration lower than 2 seconds and a total page view of 0 or 1 will be removed,
- *The graphical analysis already provides a important insight given the business objectives. From the unbalanced data available we can conclude that around of 46% shopping carts are lost on that session. That raises a question of how are this recovered (example on a next session) or if this means that all this sales are lost right at the end of the sales funnel*.

